Centrality and Activity for Envolving Networks

### 2013-06-24 network-science/paper.tex

\begin{abstract}
We address the dynamic properties of typical social networks, in which people
are considered as vertices while relations are considered as edges. We suppose
the evolution of a social network overtime only includes bidirectional new
connections between existing or newly created nodes, and also assume the graph
will evolve as a typical social network. We study the activity and centrality
metrics for nodes in such graphs by regulating a series of reasonable
properties the metrics should satisfy. We then formally define our metrics and
give their mathematical analysis. We have developed a practical based framework
to maintain the evolving large-scale network and the algorithm to compute our
metrics in preferable time and space consumption, including historical queries.
Finally, we use a simple model to generate data and apply our results to such
experiments.\\[1.5em]

@@

\begin{definition}[Evolving Centrality]
\textbf{Evolving centrality} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DA^\bullet \mathbf{x} + \beta\mathbf{1},
\end{equation}
$A^\bullet$ is the 0/1 adjacency matrix of $E^\bullet_t$ as defined in section 2,
because centrality accentuate more on the structure of the network but not the
intensity of edges. And
$D$ is a diagonal matrix that
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases}
\end{equation}
Then $\textbf{x}_i(G_t)$ is the centrality for node $v_i$ in $G_t$, and is
scale-relative to the influence of $v_i$ in the network at time $t$.
\end{definition}

The equivalent form of evolving centrality $x(G_t,v_i)$ is 
\begin{equation}
x_i(G_t,v_i)=\alpha\sum_jA^\bullet_{ij}
\mathit{Act}(G_t,v_j)x_j(G_t,v_j)+\beta.
\end{equation}

@@

\begin{definition}[Evolving PageRank]
\textbf{Evolving PageRank} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DKA \mathbf{x} + \beta\mathbf{1},
\end{equation}
where $A_{ij}$ stands for the total weight of edges from $j$ to $i$,
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases};\qquad
K_{ij}=\begin{cases}
0&(i\neq j)\\
k_{\mathrm{out}}(G_t,v_i)&(i=j)
\end{cases}.
\end{equation}
\end{definition}

@@

\begin{algorithm}[htbp]
\caption{\textsc{Query Centrality}($t$,$\alpha$,$\beta$,
$\mathbf{x}_0$,$m$)}
\label{alg-cen}
\textbf{Time Complexity:} $O(mK|V|)$, $K$ is the sparsity constant.\\
\textbf{Additional Space:} $O(|V|)$
\begin{algorithmic}[1]
\Require $0<\alpha<\lambda_1(A^\bullet(G_t))^{-1}$ and $\beta>0$ and $m>0$. Here
$\lambda_1(A)$ is the maximal possitive eigenvalue of matrix $A$.
\Ensure return the active centrality $\mathbf{x}(G_t)$.
\For{$i\gets 1\dots|V|$}
\State $a[i]\gets \Call{Query Activity}{t,v_i}$;
\State $\mathbf{x}[i]\gets \mathbf{x}_0[i]$;
\EndFor
\For{$\mathit{count}\gets 1\dots m$}
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x'}[i]\gets \beta$;
\EndFor
\For{$i\gets 1\dots|V|$}
\ForAll{$j\in\mathit{edgeListReduced}[i]$}
\State $\mathbf{x'}[i]\gets \mathbf{x'}[i]+\alpha\cdot a[j]\cdot\mathbf{x}[j]$
\EndFor\EndFor
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x}[i]\gets \mathbf{x'}[i]$;
\EndFor
\EndFor
\State\Return{$\mathbf{x}$};
\end{algorithmic}
\end{algorithm}

\subsection{Optimization and Parallelizability}

In most cases, we only need to query the latest properties of the graph, i.e.,
$t=\mathit{currentTime}$. And we observe that in Algorithm \ref{alg-con},
\ref{alg-deg} and \ref{alg-act}, the binary search can be replace by simply find
the last entry of the argument sequence. So those three algorithms can be done
in $O(1)$ time if we do not perform historical queries.

Algorithm \ref{alg-cen} actually repeat the computation of the multiplication
of a sparse matrix to a vector, which is highly parallelizable[?]. So in a large
scale distributed system, we can perform very efficient analysis to an evolving
graph, by our model.
\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunatly, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that there will always
be just one component -- known as the giant component. Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced another way to generate
artificial data of evolving social networks\cite{kumar2010structure}.

@@

\subsection{Com}

\section{Conclusion and Perspective}
% TODO

>>>>

\begin{abstract}
We address the dynamic properties of typical social networks, in which people
are considered as vertices while relations are considered as edges. We suppose
the evolution of a social network overtime includes new
connections between existing or newly created nodes, and also assume the graph
will evolve as a typical social network. We study the activity and centrality
metrics for nodes in such graphs by regulating a series of reasonable
properties the metrics should satisfy. We then formally define our metrics and
give their mathematical analysis. We have developed a practical based framework
to maintain the evolving large-scale network and the algorithm to compute our
metrics including historical queries in preferable time and space consumption.
Finally, we use a simple model to generate evolving data from existing social
network data and compare our metrics to static metrics. The results could imply
that dynamic centrality is more capable in spotting hot social events.\\[1.5em]

@@

\begin{definition}[Evolving Centrality]
\textbf{Evolving centrality} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DLA^\bullet \mathbf{x} + \beta\mathbf{1},
\end{equation}
$A^\bullet$ is the 0/1 adjacency matrix of $E^\bullet_t$ as defined in section 2,
because centrality accentuate more on the structure of the network but not the
intensity of edges. And
$D,L$ are diagonal matrices such that
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases}\qquad
L_{ij}=\begin{cases}
0&(i\neq j)\\
1/k^\bullet(G_t,v_i)&(i=j)
\end{cases}.
\end{equation}
Then $\textbf{x}_i(G_t)$ is the centrality for node $v_i$ in $G_t$, and is
scale-relative to the influence of $v_i$ in the network at time $t$.
\end{definition}

The equivalent form of evolving centrality $x(G_t,v_i)$ is 
\begin{equation}
x_i(G_t,v_i)=\alpha\sum_j\frac{A^\bullet_{ij}
\mathit{Act}(G_t,v_j)x_j(G_t,v_j)}{k^\bullet(G_t,v_j)}+\beta.
\end{equation}

@@

\begin{definition}[Evolving PageRank]
\textbf{Evolving PageRank} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DLA \mathbf{x} + \beta\mathbf{1},
\end{equation}
where $A_{ij}$ stands for the total weight of edges from $j$ to $i$,
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases};\qquad
L_{ij}=\begin{cases}
0&(i\neq j)\\
1/k_{\mathrm{out}}(G_t,v_i)&(i=j)
\end{cases}.
\end{equation}
\end{definition}

@@

\begin{algorithm}[htbp]
\caption{\textsc{Query Centrality}($t$,$\alpha$,$\beta$,
$\mathbf{x}_0$,$m$)}
\label{alg-cen}
\textbf{Time Complexity:} $O(mK|V|)$, $K$ is the sparsity constant.\\
\textbf{Additional Space:} $O(|V|)$
\begin{algorithmic}[1]
\Require $0<\alpha<\lambda_1(A^\bullet(G_t))^{-1}$ and $\beta>0$ and $m>0$. Here
$\lambda_1(A)$ is the maximal possitive eigenvalue of matrix $A$.
\Ensure return the active centrality $\mathbf{x}(G_t)$.
\For{$i\gets 1\dots|V|$}
\State $a[i]\gets \Call{Query Activity}{t,v_i}$;
\State $d[i]\gets \mathrm{size}(\mathit{edgeListReduced}[i])$;
\State $\mathbf{x}[i]\gets \mathbf{x}_0[i]$;
\EndFor
\For{$\mathit{count}\gets 1\dots m$}
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x'}[i]\gets \beta$;
\EndFor
\For{$i\gets 1\dots|V|$}
\ForAll{$j\in\mathit{edgeListReduced}[i]$}
\State $\mathbf{x'}[i]\gets \mathbf{x'}[i]+\alpha\cdot
a[j]\cdot\mathbf{x}[j]/d[j]$
\EndFor\EndFor
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x}[i]\gets \mathbf{x'}[i]$;
\EndFor
\EndFor
\State\Return{$\mathbf{x}$};
\end{algorithmic}
\end{algorithm}

\subsection{Optimization and Parallelization}
In most cases, we only need to query the latest properties of the graph, i.e.,
$t=\mathit{currentTime}$. And we observe that in algorithm \ref{alg-con},
\ref{alg-deg} and \ref{alg-act}, the binary search can be replace by simply find
the last entry of the argument sequence. So those three algorithms can be done
in $O(1)$ time if we do not perform historical queries.

What the core part of algorithm \ref{alg-cen} does is only the multiplication
of a sparse matrix to a vector, which is a classical problem called sparse
matrix-vector multiplication (SpMV). This problem can be easily solve in both
throughput-oriented and latency-oriented processors in
parallel\cite{williams2009optimization, bell2008efficient,
bell2009implementing, catalyurek1999hypergraph, zhuo2005sparse}.  This is a
good news to us because nowadays parallel machine is much more common in
companies, research facilities and even with desktop computers, thanks to the
emergence of GPGPU platforms such as CUDA and OpenCL.

\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunately, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that there will always
be just one component -- known as the giant component. Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced another way to generate
artificial data of evolving social networks\cite{kumar2010structure}.

@@

The time efficiency in terms of $|V|$ proves to have logarithm growth as
expected.

\subsection{Centrality Analysis}
Here are some traces of both the evolving centrality and the classical PageRank
centrality of specific nodes.

We generates evolving network data in the same model discussed above, with
$|V|\leq 6\,000$ and $|E|\leq 30\,000$. All iterations run for enough time,
until the maximum magnitude of change in individual centralities (known as the
$\infty$-norm of $\Delta\mathbf{x}$) is smaller than $10^{-4}$ by absolute value.
We also calculate the static PageRank centrality in parallel, which do not have
the activity metric as coefficients. All of our calculation assumes
$\alpha=0.8$, and $\beta=1$.

The centrality trace graph for an representative node, when $s$:

\section{Conclusion and Perspective}
So far, we have established and verified the framework for analysing an
evolving social network. We defined two metrics for the networks, namely
\textbf{evolving activity} and \textbf{evolving centrality}. The
experiment showed that our model can successful compute both of the metrics in
practical cases.

By the final experiment, we can observe that the increase on magnitude of
evolving centrality is similar to that of the static one. However, when the
influence of a node stops increase so fast, our evolving centrality quickly
pulls back, while the static centrality usually keeps a high value for long
time. In many fields, the social concern are supposed to be sensitive to time
, innovation and activeness, instead of imposing more attention to historical
issues that are once very popular. Those fields such as web-page search
engine[?], ranking of movie actors[?], seeking hot scholar topics, etc.
By using evolving centrality, the increment of actions of a node and its
closely related region will be more notable in increment of its ranking of
influence, whereas the static centrality is more likely to trigger cascade on
social preference[?] because people will have less chance to see and rate new
information. The cascade may prevent high quality information from being
popular, and even worse, hasten the monopolization in the commercial aspect.

The evolving centrality is more sensitive to activities of the surroundings, so
it usually changes more dramatically than static centrality which is more
stable. Dynamic metrics provides new methods for network analysis.
